Alleviating the Topology Mismatch Problem in Unstructured and Structured P2P Overlay Networks: A Survey

### 2014-01-23 unstructured.tex
2012.07.12 (Vassilis) Merged introduction section. Sectioning according to the method an algorithm is using dropped (almost all use a combination of two or more). First attempt for the tabular data on an external file.

\section{Decentralized Unstructured Algorithms}
\label{section:unstructured}

This section is dedicated to the unstructured P2P architectures. In the next
paragraphs,algorithms for tackling the topology mismatch problem on such
networks are visited, analysed and categorised based on the use of the overlay
structure, messaging approach for peer communication and techniques proposed for
detecting proximity in order to optimise overlay topology. In short, the methods
that are highlighted here are divided into
\begin{inparaenum}[\itshape i\upshape)]
  \item broadcast-based,
  \item cache-based,
  \item overlay optimisation-based, and
  \item landmark-based
\end{inparaenum}
 approaches.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Broadcast Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the first realisation of the Gnutella protocol, each query request received
by a peer is then forwarded to all its neighbours, clearly generating
unnecessary traffic over the network. The algorithms highlighted in this
paragraph are referred to as broadcast-based because they enhance the "naive"
broadcasting approach described above and prefer forwarding messages to a
selected subset of a node's outgoing links. The selection criteria vary
depending on the algorithm but they mainly use one or multiple statistical
metrics that span from the obvious ones, like the latency of the link between
the nodes, to more sophisticated alternatives (that may even take into account
application level requirements) such as the reliability of the neighbour. The
forwarding-based approach has the advantage of enhancing search efficiency but
on the other hand suffer drawbacks like the drastic reduction of the search
scope because of the restrictions of the broadcasting policies as well as
limitations when scaling to large networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Improving search in peer-to-peer networks}

@@

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency problem caused by blind flooding in Gnutella-like
P2P file sharing protocols. The paper replaces blind flooding with three
approaches, namely \textit{iterative deepening}, \textit{directed BFS}, and
\textit{local indices}. In \textit{iterative deepening}, as shown in
Figure~\ref{figure:dbfs:iterdeep}, the search is performed on a BFS tree with
multiple preset depths. The depth limit is iteratively increased by the source
node for each query based on the quality of the results. The source node, after
examining the results, may issue a new request by increasing the depth limit,
which will trigger the nodes at the last depth level to resume the search. The
iterative approach avoids to start the search from scratch at each iteration and
reduces the load of the nodes on the upper levels of the tree. The major
drawback of \textit{iterative deepening} is the delay between successive
iterations, as the source node needs to examine the results at each iteration
before deciding to quit or resume the query.  The \textit{directed BFS}, as
shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this delay by
forwarding the query messages only to a selected set of neighbours, in which the
selection criteria varies from the number of results received previously,
distance in terms of hops, bandwidth, or the query load of the neighbour. In
\textit{local indices}, as shown in Figure~\ref{figure:dbfs:localindx}, each
node sustains a local data index of the nodes within a radius of $r$ hops from
itself and uses this local index to remotely query the neighbour nodes without
generating additional traffic. \textit{Local indices} greatly reduces the
aggregate bandwidth usage of the network and improves query efficiency, however,
updating the indices in cases with frequent node joins and leaves introduces a
serious overhead to the system if the radius is kept broad.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Gia}
\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
problem of the unstructured P2P file sharing systems, in particular Gnutella.
The main novelty in the design of Gia is the replacement of the blind flooding
approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
random walks is a step in the right direction, issuing only a single copy of the
query within the network reduces the search scope, thus affecting negatively the
success rate of the query.  In order to overcome this limitation, Gia introduces
a token-based flow control mechanism, which is essentially an intelligent flow
control algorithm that gradually redirects the queries to nodes which are more
likely to answer. In order to prevent overloading of nodes with query requests,
Gia uses a token-based flow control algorithm in which each node announces the
number of query requests it can handle in terms of tokens to its neighbours, so
that neighbours only forward query requests to nodes that they previously
received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
processing power, disk speed, etc, of the nodes in P2P networks and uses this
information when connecting nodes to each other and by using a topology
adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
low capacity nodes are within short proximity to high capacity ones.
% TODO: THIS FOLLOWING REMARK SOUNDS LIKE THE ALGORITHM SHOULDN'T BE IN THIS
%       SURVEY AFTER ALL!!
Although the topology adaptation algorithm Gia uses improves the scalability of
the network, it does not help much in solving the topology mismatch problem,
since it does not consider the underlying physical topology.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Distributed Cycle Minimization Protocol}
We have already discussed that blind flooding mechanism broadcasts messages with
no real targeting, many times to a direction far from the path that can
ultimatelly fullfil the query request. Another problem with topology unaware
systems is the duplication of messages due to cycles, even along the correct
forwarding path. The authors of \cite{zhu_dcmp_2008} fucused on that exact
deficiency of overlay networks and introduced Distributed Cycle Minimization
Protocol (DCMP), a dynamic, fully distributed method to remove cycles without
sacrificing overlay connectivity, resilience and other important properties of
unstructured architectures by avoiding a hierarchical organisation of peers. In
DCMP, once a cycle is detected, the most powerful node in that cycle is
elected as the \emph{Gate Peer} and the cycle is then broken at a strategic
place so that it will result in the minimisation of the distance between the
Gate Peer and all other nodes that participated in that cycle. The process
is managed by using two special message types namely \emph{Information
Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage of DCMP is that
since the distance a forwarded message can travel is limited by the TTL value,
which is practically $7$ in most cases DCMP cannot detect cycles formed by more
than 7 peers. Even though cycle elimination improves the network performance, it
does not actually solve the topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cache Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Replication Strategies in Unstructured Peer-to-Peer Networks}
\cite{Cohen02} aim to improve the inefficient blind search algorithm by
replicating the data in a P2P network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. In order to analyse the feasibility of
such an approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed in this paper, which is a model between the
uniform and proportional allocation.  In order to evaluate the outcome of the
replication models, the overall costs of successful and unsuccessful searches in
the network are compared. Results argue that the uniform allocation model
minimizes the maximum search size, therefore reduces the time spent on an
unsuccessful search. The proportional model, on the other hand, promotes the
more frequently queried items by replicating them more, therefore decreases the
search time for popular items, but suffers when needing to locate the rare
items. The authors also claim that the expected successful search size is the
same for uniform and proportional models, and any approach between them would
behave much better. Therefore they propose the square root allocation approach,
which is a replication strategy that minimizes the expected search size of
successful queries in P2P networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Tracing a large-scale Peer to Peer System: an hour in the life
of Gnutella}
\cite{Markatos02} analyses Gnutella network traffic traces and by concluding
there is locality among query requests, proposes a caching algorithm that tries
to exploit these findings . The analysis of the trace data reveals other
important facts about the structure of the Gnutella network and the query data.
One significant observation is that the geographic locations of clients do not
have a correlation with the number of query requests they receive. This is an
obvious result of the topology mismatch problem caused by the overlay structure
of the Gnutella network. Gnutella's traffic is observed to be bursty both for
query requests and query responses, even in longer intervals. It is observed
that a peer receives 50 query messages per second on average! Moreover nine out
of ten queries do not generate any response due to the inefficient design of
the Gnutella network. When developing a caching system to exploit locality,
applying an approach similar to web caching does not fit well with P2P systems.
Caches in P2P systems not only have to consider the query string, but also
the TTL value, the source of the query, and the time of the query as well.  In
general, even though optimum caching is hard to achieve, it is reported that
it improves the overall performance of the Gnutella network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Overlay Optimization Based}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Narada}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Adaptive Overlay Topology Optimization and Adaptive Connection
Establishment}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_aoto_2003} is one
of the first attempts, along with Narada, to address the topology mismatch
problem. AOTO is a distributed algorithm that works in two steps, namelly
\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
the underlying physical resourses. In \emph{Selective Flooding} a minimum
spanning tree is built among each peer and its immediate neighbours so that
queries are not flooding the whole network and at the same time without
shrinking the search scope. During \emph{Active Topology}, the overlay
links are revised so that they depict the physical network topology as close as
possible. This is done independently, by each peer by replacing non-flooding
neighbours, with closer nodes. Picking a replacement out of these non-fooding
neighbours follows a random policy (called \emph{Randomized AT} algorithm in the
paper's context). For these steps a peer needs to keep track of communication
costs to all its neighbours (e.g. the network delay) as well as between any
pair of neighbours (meaning also that additional, cost probing, message types
need to be added to the Gnutella protocol). Whenever a new neighbour cost table
is received or there is a change of neighbours, the source peer has to
re-calculate the multicast tree and apply the randomized AT algorithm.

%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbours. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbours.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbours has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbour replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.

%
% TODO: REMAKE AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA!!!!!!!!!!!!!!!!!!
%
 In
\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the authors
extend the idea by introducing optimizations based on the depths of the minimum
spanning trees. But since the network delay is not always a reliable estimation
method to detect the physical location of peers, the algorithm still suffers
from the discrepancies caused by mislocated nodes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Location-aware Topology Matching}
\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
method to optimize the overlay structure of the P2P network based on the
physical topology. To achieve this, peers issue special messages called
\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
them and calculating communication costs. Calculation is achieved through
timestamp checking, so clocks must be synchronised. The latency information
gathered is later used to evolve the overlay network into a more efficient one,
without reducing the search scope. Each node examines the latency information
among his direct neighbours and those with larger latency are put in to a
will-cut list where they remain for a certain period of time until when they are
ultimatelly cut and recorded to the peer's cut list. Thus, low productive
connections are dropped and replaced by more efficient ones, reducing the
latency on the overall network. Although LTM improves the overall efficiency of
the P2P network, it does not actually use real physical topology information,
just latency metrics, therefore it does not offer a guaranteed safety net for
the topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Scalable Bipartite Overlay}
\emph{Scalable Bipartite Overlay (SBO)}
\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of
creating and maintaining a minimum spanning tree cost by randomly dividing the
nodes into two groups, namely red and white and assigning different tasks to
different groups. When a peer joins the network, is randomly assigned with an
initial colour, say red or white (separating all peers into two groups). Then
the peer that plays the role of the bbootstrap host provides the joining peer
with a list of active peers along with their colour information. The later uses
this list to establish connections to differently coloured peers. This way, all
peers form a bipartite overlay. The white peers measure distances to
neighbours (red only) by using the network delay as a metric and report their
corresponding red neighbours. The red peers, equipped with the information of
all two-hop neighbours ($N^2$), creates minimum spanning tree of these
neighbours and assigns efficient forwarding paths. Having a minimum spanning
tree with two hops, a red peer is able to send its queries within that range.
Some white peers, though, have become non-forwarding neighbours. In this phase,
such a (white) neighbour will try to find another red peer being two hops away
from its current red neighbour to replace the later as its new neighbour. The
white peers can further optimize their positions within the overlay if this is
required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
\paragraph*{\bf Adaptive Peer Selection}
Bernstein et al., contributed on the use of machine learning for building peer
selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
decision tree is used to rate peers based on information about connection
characteristics that is collected. These can be load, bandwidth, and past
uploading experience. Then a by Markov decision process is incorporated as a
mechanism to shape the policy for switching among the peers. The problem
with this approach is that the peer selection algorithm is slow due to the
learning process and the complexity of the method.

>>>>

\section{Decentralized Unstructured Architectures}
\label{section:unstructured}

This section is dedicated to the unstructured P2P architectures. In the next
paragraphs,algorithms for tackling the topology mismatch problem on such
networks are discussed, analysed and categorised based on the use of the
overlay structure, message forwarding approach for peer communication and
techniques proposed for detecting proximity in order to optimise overlay
topology.

@@

\subsection{Algorithms for Unstructured Architectures}

%###############################################################################
%###############################################################################
%       BROADCAST OPTIMISATION
%###############################################################################
%###############################################################################

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Improving search in peer-to-peer networks}

@@

\cite{yang_improvep2psearch_2002} proposes an easy to implement and practical
solution to the inefficiency of Gnutella's ``blind flooding'' approach. The
paper introduces three different forwarding methods, namely \emph{iterative
deepening}, \emph{directed BFS}, and \emph{local indices}.

In iterative deepening (see Figure~\ref{figure:dbfs:iterdeep}), the search is
performed on a BFS tree with multiple preset depths. The depth limit is
iteratively increased by the source node for each query, based on the quality of
the results. The source node may issue a new request with increased depth limit
that will trigger the nodes at the last depth level to resume the search. The
iterative approach avoids restarting the whole search process from scratch at
each iteration and reduces the load of the nodes on the upper levels of the
tree. Its major drawback is the delay between successive iterations, as the
source node needs to examine the results at each iteration before deciding to
quit or resume the query.

The directed BFS, as shown in Figure~\ref{figure:dbfs:dbfs}, tries to avoid this
delay by forwarding the query messages only to a selected set of neighbours, in
which the selection criteria varies from the number of results received
previously, distance in terms of hops, bandwidth, or the query load of the
neighbour.

% TODO: local indices is a cache based algorithm so maybe we have to put a
% reference to the respective section???
In local indices, as shown in Figure~\ref{figure:dbfs:localindx}, each node
indexes data within a radius of $r$ hops and uses this local index to answer
queries on behalf of them without generating additional traffic. Local indices
greatly reduce the aggregate bandwidth usage of the network and improves query
efficiency, however, updating the indices in cases with frequent node joins and
leaves introduces a serious overhead to the system if the radius is kept broad.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Gia}
\emph{Gia} \cite{chawathe_gia_2003} is an effort to solve the scalability
problem of the unstructured P2P file sharing systems, in particular Gnutella.
The main novelty in the design of Gia is the replacement of the blind flooding
approach of Gnutella with random walks \cite{lv_randomwalks_2002}. Although
random walks is a step in the right direction, issuing only a single copy of the
query within the network reduces the search scope, thus affecting negatively the
success rate of the query.  In order to overcome this limitation, Gia introduces
a token-based flow control mechanism, which is essentially an intelligent flow
control algorithm that gradually redirects the queries to nodes which are more
likely to answer. In order to prevent overloading of nodes with query requests,
Gia uses a token-based flow control algorithm in which each node announces the
number of query requests it can handle in terms of tokens to its neighbours, so
that neighbours only forward query requests to nodes that they previously
received tokens from. Gia also acknowledges the heterogenity in peer bandwidth,
processing power, disk speed, etc, of the nodes in P2P networks and uses this
information when connecting nodes to each other and by using a topology
adaptation algorithm, Gia ensures that high capacity nodes have high degrees and
low capacity nodes are within short proximity to high capacity ones.
% TODO: THIS FOLLOWING REMARK SOUNDS LIKE THE ALGORITHM SHOULDN'T BE IN THIS
%       SURVEY AFTER ALL!!
Although the topology adaptation algorithm Gia uses improves the scalability of
the network, it does not help much in solving the topology mismatch problem,
since it does not consider the underlying physical topology.

% TODO: from the following (commented out) design components of Gia has one
% part that falls into topology adaptation, one part on forwarding optimisation
% and one part on

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Distributed Cycle Minimization Protocol (DCMP)}
We have already discussed that blind flooding mechanism broadcasts messages with
no real targeting, many times to a direction far from the path that can
ultimatelly fullfil the query request. Another problem with topology unaware
systems is the duplication of messages due to cycles, even along the correct
forwarding path. The authors of \cite{zhu_dcmp_2008} fucused on that exact
deficiency of overlay networks and introduced Distributed Cycle Minimization
Protocol (DCMP), a dynamic, fully distributed method to remove cycles without
sacrificing overlay connectivity, resilience and other important properties of
unstructured architectures by avoiding a hierarchical organisation of peers. In
DCMP, once a cycle is detected, the most powerful node in that cycle is
elected as the \emph{Gate Peer} and the cycle is then broken at a strategic
place so that it will result in the minimisation of the distance between the
Gate Peer and all other nodes that participated in that cycle. The process
is managed by using two special message types namely \emph{Information
Collection (IC)} and \emph{Cut Message (CM)}. One disadvantage of DCMP is that
since the distance a forwarded message can travel is limited by the TTL value,
which is practically $7$ in most cases DCMP cannot detect cycles formed by more
than 7 peers. Even though cycle elimination improves the network performance, it
does not actually solve the topology mismatch problem.

@@

%###############################################################################
%###############################################################################
%###############################################################################
%       CACHING
%###############################################################################
%###############################################################################
%###############################################################################

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Replication Strategies in Unstructured Peer-to-Peer Networks}
\cite{Cohen02} aim to improve the inefficient blind search algorithm by
replicating the data in a P2P network. The main intuition behind the
idea of replication, or using cached copies, is that as the number of copies for
each item increases in the network, it would be easier for a search algorithm,
even a random one, to locate these items. In order to analyse the feasibility of
such an approach, the authors investigate three different replication
strategies, namely uniform, proportional, and square root allocation. In the
uniform model, the copies of items are uniformly replicated in the network,
while in the proportional model, the items are replicated based on their query
rate, so that frequently queried items are replicated more. The square root
allocation is a strategy proposed in this paper, which is a model between the
uniform and proportional allocation.  In order to evaluate the outcome of the
replication models, the overall costs of successful and unsuccessful searches in
the network are compared. Results argue that the uniform allocation model
minimizes the maximum search size, therefore reduces the time spent on an
unsuccessful search. The proportional model, on the other hand, promotes the
more frequently queried items by replicating them more, therefore decreases the
search time for popular items, but suffers when needing to locate the rare
items. The authors also claim that the expected successful search size is the
same for uniform and proportional models, and any approach between them would
behave much better. Therefore they propose the square root allocation approach,
which is a replication strategy that minimizes the expected search size of
successful queries in P2P networks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Tracing a large-scale Peer to Peer System: an hour in the
life
of Gnutella}
% TODO: This looks better as an analysis for the unstructured networks in
% general and specifically for caching, so it doesn't seem to contribute any
% new approach.... At least we do not present anything here. Maybe we need to
% revisit the paper itself once more
\cite{Markatos02} analyses Gnutella network traffic traces and by concluding
there is locality among query requests, proposes a caching algorithm that tries
to exploit these findings . The analysis of the trace data reveals other
important facts about the structure of the Gnutella network and the query data.
One significant observation is that the geographic locations of clients do not
have a correlation with the number of query requests they receive. This is an
obvious result of the topology mismatch problem caused by the overlay structure
of the Gnutella network. Gnutella's traffic is observed to be bursty both for
query requests and query responses, even in longer intervals. It is observed
that a peer receives 50 query messages per second on average! Moreover nine out
of ten queries do not generate any response due to the inefficient design of
the Gnutella network. When developing a caching system to exploit locality,
applying an approach similar to web caching does not fit well with P2P systems.
Caches in P2P systems not only have to consider the query string, but also
the TTL value, the source of the query, and the time of the query as well.  In
general, even though optimum caching is hard to achieve, it is reported that
it improves the overall performance of the Gnutella network.

%###############################################################################
%###############################################################################
%###############################################################################
%       OVERLAY OPTIMISATION
%###############################################################################
%###############################################################################
%###############################################################################

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Narada}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Overlay Topology Optimization (AOTO)}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_aoto_2003} is one
of the first attempts, along with Narada, to address the topology mismatch
problem. AOTO is a distributed algorithm that works in two steps, namelly
\emph{Selective Flooding} and \emph{Active Topology} to optimize the usage of
the underlying physical resourses. In \emph{Selective Flooding} a minimum
spanning tree is built among each peer and its immediate neighbours so that
queries are not flooding the whole network and at the same time without
shrinking the search scope. During \emph{Active Topology}, the overlay
links are revised so that they depict the physical network topology as close as
possible. This is done independently, by each peer by replacing non-flooding
neighbours, with closer nodes. Picking a replacement out of these non-fooding
neighbours follows a random policy (called \emph{Randomized AT} algorithm in the
paper's context). For these steps a peer needs to keep track of communication
costs to all its neighbours (e.g. the network delay) as well as between any
pair of neighbours (meaning also that additional, cost probing, message types
need to be added to the Gnutella protocol). Whenever a new neighbour cost table
is received or there is a change of neighbours, the source peer has to
re-calculate the multicast tree and apply the randomized AT algorithm.

%\paragraph*{Selective Flooding (SF)}
%
% TODO: I DON'T REMEMBER WHY THE FOLLOWING IS MENTIONING LTM WHICH IS ANOTHER
%       ALGORITM. MAYBE THIS CAN BE USED AS A PART OF DISCUSSION AND/OR
%       COMPARISON OF AOTO AND LTM
% TODO: EDIT: LTM SEEMS A MISTAKE HERE BETTER NOT USE IT FOR DISCUSSION (THIS
%             IS AOTO)
%LTM's SF effectiveness has be proven to be detached from the different physical
% or overlay topologies. On the other hand, SF is more effective with large
%number of logical neighbours. It can reach an average optimization rate of 87.4
%percent on a logical topology with an average of 30 logical neighbours.
%
%\paragraph*{Active Topology (AT)}
%
%Different numbers of average logical neighbours has little to do with the
% effectiveness of AT. If the source has $n$ non-flooding peers, there are $n$
%potential neighbour replacements. The overhead to exhaust all $n$ possible
%replacements can be high, so in practice, after each replacement the source
%peer can decide whether it needs to find another candidate peer. This is done
%by computing the cost improvement ratio greater than some predefined
%termination threshold. The larger the threshold, the slower, in the number of
%optimization steps, the reduction of the normalized average distance. As a
%whole the average response time is significantly reduced when more optimization
%steps taken.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Adaptive Connection Establishment (ACE)}
\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an
overlay multicast tree among each source node and the peers within a certain
diameter from the source peer and optimizes the neighbour connections that are
not in that tree. To achieve that it calculate cost between nodes using network
delay as a metric. Each peer probes the costs with its immediate logical
neighbours and forms a \emph{neighbour cost table (NCT)} using a special
routing message type. Two neighbouring peers exchange their NCTs in order for
every peer to obtain the cost between any pair of its local neighbours forming a
small overlay topology. Moreover, based on obtained NCTs a minimum spanning tree
among each peer and its immediate neighbours is built. Ultimately, physically
far away neighbours are replaced by physically close ones. Specifically, a peer
$S$ probes the distance between one of its non-flooding neighbour's neighbour,
say $G$ and $H$ respectively. Assuming that the link to the neighbour's neigbour
($SH$) is smaller that that of the neighbour ($SG$), the later is cut and a new
connection is established. In the oposite situation where $SG$ is smaller than
$SH$ then if $SH$ is smaller that $GH$ then $H$ is preserved as $S$'s new
neighbour. In the case where $SH$ is larger than $SG$ and $GH$, no connection
will be established and $S$ will continue probing another neighbour's neighbour.
The above optimisation is conducted within $1$-neighbour closure (among its
source peer and all its direct neighbours) but the scope can be extended. The
larger the scope, the better topology matching improvement but also the greater
the computational overhead.

%In \emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004}, the
%authors extend the idea of AOTO by introducing optimizations based on the
%depths of the minimum spanning trees. But since the network delay is not
%always a reliable estimation method to detect the physical location of peers,
%the algorithm still suffers from the discrepancies caused by mislocated nodes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Location-aware Topology Matching (LTM)}
\emph{Location-aware Topology Matching (LTM)} \cite{liu_ltm_2004} proposes a
method to optimize the overlay structure of the P2P network based on the
physical topology. To achieve this, peers issue special messages called
\textit{TTL-detector}s with TTL values of 0 and 1 so that the receiving peers
discover one- and two-hop neighbour sets ($N$ and $N^2$ respectively) around
them and calculating communication costs. Calculation is achieved through
timestamp checking, so clocks must be synchronised. The latency information
gathered is later used to evolve the overlay network into a more efficient one,
without reducing the search scope. Each node examines the latency information
among his direct neighbours and those with larger latency are put in to a
will-cut list where they remain for a certain period of time until when they are
ultimatelly cut and recorded to the peer's cut list. Thus, low productive
connections are dropped and replaced by more efficient ones, reducing the
latency on the overall network. Although LTM improves the overall efficiency of
the P2P network, it does not actually use real physical topology information,
just latency metrics, therefore it does not offer a guaranteed safety net for
the topology mismatch problem.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Scalable Bipartite Overlay (SBO)}
\emph{Scalable Bipartite Overlay (SBO)}
\cite{liu_bipartite_IPDPS,liu_bipartite_2007} reduces the overhead of creating
and maintaining a minimum spanning tree cost by randomly dividing the nodes into
two groups, namely red and white and assigning different tasks to different
groups. When a peer joins the network, is randomly assigned with an initial
colour, say red or white (separating all peers into two groups). Then the peer
that plays the role of the bbootstrap host provides the joining peer with a list
of active peers along with their colour information. The later uses this list to
establish connections to differently coloured peers. This way, all peers form a
bipartite overlay. The white peers measure distances to neighbours (red only) by
using the network delay as a metric and report their corresponding red
neighbours. The red peers, equipped with the information of all two-hop
neighbours ($N^2$), creates minimum spanning tree of these neighbours and
assigns efficient forwarding paths. Having a minimum spanning tree with two
hops, a red peer is able to send its queries within that range. Some white
peers, though, have become non-forwarding neighbours. In this phase, such a
(white) neighbour will try to find another red peer being two hops away from its
current red neighbour to replace the later as its new neighbour. The white peers
can further optimize their positions within the overlay if this is required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: DOUBLE CHECK IF THIS ALGORITHM GOES IN THIS SECTION/SUBSECTION
\subsubsection{Adaptive Peer Selection}
Bernstein et al., contributed on the use of machine learning for building peer
selection strategies from past experience \cite{bflz_adaptpeersel_2003}. A
decision tree is used to rate peers based on information about connection
characteristics that is collected. These can be load, bandwidth, and past
uploading experience. Then a by Markov decision process is incorporated as a
mechanism to shape the policy for switching among the peers. The problem
with this approach is that the peer selection algorithm is slow due to the
learning process and the complexity of the method.
